SMT-Boosted Security Types for Low-Level MPC

### 2025-05-05 esop25/slides.tex
checkpoint

  \cemph{Secure Multiparty Computation (MPC)} is a paradigm for
  \cemph{confidential computing}.
  \begin{itemize}
  \item Supports data privacy.
  \item Uses software-based protocols (vs., e.g., enclaves).
  \item No trusted 3rd party.
  \end{itemize}
  Many existing deployments: voting, auctions, COVID exposure notification, federated deep learning...\footnote{https://mpc.cs.berkeley.edu}

@@

  Existing high-level PLs rely on low-level MPC protocols with correctness
  assumed.
  \begin{itemize}
  \item Distinct features, e.g., probabilistic sampling.
  \item Existing proof techniques mostly manual.
  \end{itemize}
  Our contributions:
  \begin{itemize}
  \item Language design for low-level protocol definitions.
  \item Automated type analysis for security properties, with correctness based on
    \cemph{hyperproperties}\footnote{Skalka and Near, PPDP 2024}.
  \item Our type analysis leverages an \cemph{SMT finite fields theory}\footnote{Ozdemir et. al, CAV 2023} for:
    \begin{itemize}
    \item \cemph{Correctness}
    \item \cemph{Confidentiality}
    \item \cemph{Integrity}
    \end{itemize}
  \end{itemize}

@@

\begin{frame}{What's an MPC Protocol?}
  A \emph{protocol} $\prog$ implements an \emph{ideal functionality} $\idealf$,
  defined in terms of operations $(+,-,*)$ in a prime field $\mathbb{F}_p$.
  \begin{itemize}
  \item The inputs are \cemph{secret values} of \cemph{clients} in a \emph{federation}
    $\setit{1,...,j}$.
  \item Clients are distributed and communicate via \cemph{message passing}.
  \item Any client can be corrupted by the adversary, who can then see their messages.
  \item The \cemph{output} $=\idealf$ is made \emph{publicly available} by $\prog$.
  \end{itemize}
  \cemph{$\idealf$ may expose some information to an adversary} trying to
  guess secret values, since it depends on inputs.

@@

  \cemph{Confidentiality} means that the protocol doesn't enable the adversary
  to make better guesses than $\idealf$.

@@

\begin{frame}{cvc5 Finite Fields Theory}

  The cvc5 Finite Fields Theory supports automated satisfiability checking of equational
  constraint definitions $\eqs$ over prime fields $\mathbb{F}_{p}$.

@@

  Overture protocols $\prog$ have a straightforward interpretation as SMFF constraints $\toeq{\prog}$ , e.g.,
  the first two lines of $\prog_{\mathit{sum3}}$ translate as:
  $$
  %\toeq{
  %  \begin{array}{c}
  %  \begin{array}{lll}
  %    \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
  %    \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} 
  %  \end{array}\\
  %  = \\
    \begin{array}{lll}
      \elab{\mesg{s1}}{2} &\eop& \elab{\secret{1}}{1} \fminus \elab{\locflip}{1} \fminus \elab{\flip{x}}{1}\quad \wedge \\ 
      \elab{\mesg{s1}}{3} &\eop& \elab{\flip{x}}{1} 
    \end{array}
    %\end{array}
    %}
 $$

 Verifiable \emph{entailment} $\eqs_1 \models \eqs_2$ means that any model of $\eqs_1$ is a model
 of $\eqs_2$. Models of $\prog$ correspond exactly to valid runs of $\prog$, therefore...

@@

  ... entailment is a mechanism for verifying protocol properties. If an interpretation
  of a protocol entails a property, it holds in all protocol runs. 
$$
\begin{array}{c}
  \toeq{\sumthree} \wedge \ttt{out} \eop \mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{3}\\ \models\\
  \ttt{out} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
\end{array}
$$

@@

  Verifying equational properties also supports verification of confidentiality.
  It is well-known that expressions of the form:
  $$
  (e - \rx{w}){\cid}
  $$
  are a means of encrypting $e$ using $\rx{w}{\cid}$ as a one-time-pad in any
  size field. This can be leveraged in typing. 
  $$    
  \inferrule[Encode]
  {\eqs \models \toeq{e} \eop \toeq{e'} - \rx{w}{\cid}\\
   \eqj{R}{\eqs}{e'}{\ty}}
  {\eqj{R;\{ \rx{w}{\cid} \}}{\eqs}{e}{\setit{\cty{\rx{w}{\cid}}{\ty}}}}
  $$
  \begin{itemize}
  \item The \emph{ciphertype} $\cty{\rx{w}{\cid}}{\ty}$ indicates the computation
    results in a value encrypted with $\rx{w}{\cid}$.
  \item $\ty$ is a conservative approximation of the dependencies in $e'$.
  \item The \emph{randomness environment} $R;\{ \rx{w}{\cid} \}$ enforces
    linearity.
  \end{itemize}

>>>>

  \cemph{Secure Multiparty Computation (MPC)} is a paradigm for
  \cemph{confidential computing} in distributed settings.
  \begin{itemize}
  \item Supports data privacy.
  \item Uses software-based protocols (vs., e.g., enclaves).
  \item No trusted 3rd party.
  \end{itemize}
  Many existing deployments: voting, auctions, COVID exposure notification, federated deep learning...\footnote{https://mpc.cs.berkeley.edu}

@@

  Existing high-level PLs rely on low-level MPC protocols with correctness and security
  assumed.
  \begin{itemize}
  \item Distinct low-level features, e.g., probabilistic sampling.
  \item Existing low-level proof techniques mostly manual.
  \end{itemize}
  Our contributions:
  \begin{itemize}
  \item Language design for low-level protocol definitions.
  \item Automated type analysis for security properties, with correctness based on
    \cemph{hyperproperties}\footnote{Skalka and Near, PPDP 2024}.
  \item Our type analysis leverages an \cemph{SMT finite fields theory}\footnote{Ozdemir et. al, CAV 2023} for:
    \begin{itemize}
    \item \cemph{Correctness}
    \item \cemph{Confidentiality}
    \item \cemph{Integrity}
    \end{itemize}
  \end{itemize}

@@

\begin{frame}{What's an MPC Protocol?}
  A \emph{protocol} $\prog$ implements an \emph{ideal functionality} $\idealf$,
  defined in terms of operations $(+,-,*)$ in a prime field $\mathbb{F}_p$.
  \begin{itemize}
  \item The inputs are \cemph{secret values} of \cemph{clients} in a \emph{federation}
    $\setit{1,...,j}$.
  \item Clients are distributed and communicate via \cemph{message passing}.
  \item Any client can be corrupted by the \cemph{adversary}, who can then see their messages.
  \item The \cemph{output} $=\idealf$ is made \emph{publicly available} by $\prog$.
  \end{itemize}
  \cemph{$\idealf$ may expose some information to an adversary} trying to
  guess secret values, since it depends on inputs.

@@

  \cemph{Confidentiality} means that the protocol doesn't enable the adversary
  to make better guesses than $\idealf$.\footnote{Passive adversary assumed for this talk.}

@@

\begin{frame}{SMT Finite Fields Theory}

  The SMT Finite Fields Theory supports automated satisfiability checking of equational
  constraint definitions $\eqs$ over prime fields $\mathbb{F}_{p}$.

@@

  Overture protocols $\prog$ have a straightforward interpretation as equality constraints $\toeq{\prog}$ , e.g.,
  the first two lines of $\prog_{\mathit{sum3}}$ translate as:
  $$
  %\toeq{
  %  \begin{array}{c}
  %  \begin{array}{lll}
  %    \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
  %    \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} 
  %  \end{array}\\
  %  = \\
    \begin{array}{lll}
      \elab{\mesg{s1}}{2} &\eop& \elab{\secret{1}}{1} \fminus \elab{\locflip}{1} \fminus \elab{\flip{x}}{1}\quad \wedge \\ 
      \elab{\mesg{s1}}{3} &\eop& \elab{\flip{x}}{1} 
    \end{array}
    %\end{array}
    %}
  $$

    Random variables (e.g., $\rx{x}{1}$) unconstrained in the interpretation.

    \cemph{Models of $\toeq{\prog}$ correspond exactly to runs of $\prog$}.

@@

    Verifiable \emph{entailment} $\eqs_1 \models \eqs_2$ means that any model of $\eqs_1$ is a model
    of $\eqs_2$.

    \medskip
    Entailment is a mechanism for verifying protocol properties. If an interpretation
    of a protocol entails a property, it holds in all protocol runs. 
$$
\begin{array}{c}
  \toeq{\sumthree} \wedge \ttt{out} \eop \mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{3}\\ \models\\
  \ttt{out} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
\end{array}
$$

@@

  Verifying equational properties also supports verification of confidentiality.
  It is well-known that expressions of the form:
  $$
  (e - \flip{w}){\cid}
  $$
  are a means of encrypting $\elab{e}{\cid}$ using $\rx{w}{\cid}$ as a one-time-pad in any
  size field. This can be leveraged in typing. 
  $$    
  \inferrule[Encode]
  {\eqs \models \toeq{e} \eop \toeq{e'} - \rx{w}{\cid}\\
   \eqj{R}{\eqs}{e'}{\ty}}
  {\eqj{R;\{ \rx{w}{\cid} \}}{\eqs}{e}{\setit{\cty{\rx{w}{\cid}}{\ty}}}}
  $$
  \begin{itemize}
  \item The \emph{ciphertype} $\cty{\rx{w}{\cid}}{\ty}$ indicates the computation
    results in a value encrypted with $\rx{w}{\cid}$.
  \item $\ty$ is a conservative approximation of the dependencies in $e'$.
  \item The \emph{randomness environment} $R;\{ \rx{w}{\cid} \}$ enforces
    linearity.
  \end{itemize}
